
ECoG Stats Report
========================================================

This report uses lmer and linear models to evaluate the statistical validity 
of the data. 

Data details:
```{r}
lockType = 'stim'; 
type     = 'power';
band     = 'hgam'; 
```

Additional setup:
-------------------------------------------------------
Load necessary libraries:
```{r}
library('lme4');
```

Set additional variables:
```{r}
dataPath = '~/Documents/ECOG/Results/Rdata/';
hem 		= 'all';
ref 		= 'nonLPCleasL1TvalCh10';
binType     = 'Bin';
byBlockStr 	= 'bySubj';

if (lockType == 'stim'){
	timeStr 	= '200msTo900ms'
}else if (lockType == 'RT'){
	timeStr 	= 'n600msTo100ms'
	}	

if (type=='erp'){  
  band 			= '';
  preFix 		= paste(binType,timeStr,hem,'ERPs',sep='')
  baseLine 		= 'sub';
  analysisType 	= 'Amp';
}else if (type=='power'){
  preFix 		= paste(binType,timeStr,hem,'ERSPs',band,sep='')
  baseLine 		= 'sub';
  analysisType 	= 'logPower';
}

fileName = paste(preFix,'Group',lockType,'Lock',baseLine,analysisType,ref,byBlockStr,sep='');
```

Load data:
```{r}
data 		<- read.csv(paste(dataPath,fileName,'.csv',sep=''),header=F)
names(data) <- c('M','blank1','T','blank2','Z','blank3','BinNum','chanId','subjId', 'blockId', 
				 'ROI','dP','HR','FAR','hem','subROI','ITC_Z','ITP_Z',
				 'Zc','ZcH','ZcCRs','ZH','ZCRs')

data$BinNum 		= as.factor(data$BinNum)
data$chanId 		= as.factor(data$chanId)
data$subjId  		= as.factor(data$subjId)
data$blockId  		= as.factor(data$blockId)
data$ROI  			= as.factor(data$ROI);
levels(data$ROI)	= c('IPS','SPL','AG')
data$chanId 		= as.factor(data$chanId)
data$hem 			= as.factor(data$hem);
levels(data$hem)	= c('left','right')
data$subROI  		= as.factor(data$subROI);
levels(data$subROI) = c('','pIPS','aIPS','pSPL','aSPL')

nBins 	= length(unique(data$BinNum));
nChans 	= length(unique(data$chanId));
nSubjs 	= length(unique(data$subjId));
nBlocks = length(unique(data$blockId));
nROIs 	= length(unique(data$ROI));
```

Simple linear models, evaluated with anova
------------------------------------------------
**For the lefts subjects:**
```{r}
subL 			= data$hem=='left'
LModel.lm 		= lm(Z ~ 1+ ROI*BinNum+subjId+chanId,data=data,subset=subL);
anova(LModel.lm)
```

**For the right subjects:**
```{r}
subR 			= data$hem=='right'
RModel.lm 		= lm(Z ~ 1+ ROI*BinNum+subjId+chanId,data=data,subset=subR);
anova(RModel.lm)
```

**Evaluate hemishphere interatction**
```{r}
FModel.lm 		= lm(Z ~ 1+ ROI*BinNum*hem+subjId+chanId,data=data);
anova(FModel.lm)
```


Mixed Models evaluated with lmer
-----------------------------------------------------

First we define the `LRT` function that takes a MER model and an index 
for a factor, and returns the Chi square statistic based on the log-likelihood 
of the full and nested models.
```{r}
LRT <- function(theModel, idx){
	modFull 	 = theModel
	modelTerms_h <-terms(theModel)
	modelTerms   = attr(modelTerms_h, "term.labels")
	updateStr    <-paste(".~.-",as.character(modelTerms[idx]))
	modNested    = update(theModel, updateStr)
	
	llFull   = logLik(modFull)
	llNested = logLik(modNested)
	
	chsqval <- 2*(llFull[1] - llNested[1])
	dfDiff  <- attr(llFull,"df") - attr(llNested,"df")	
	pVal    <- 1-pchisq(chsqval,dfDiff)
	
	out <- c(as.character(modelTerms[idx]),pVal, chsqval, dfDiff)
	names(out) <- c('Factor', ' P val ' , ' ChiSq ', ' dF ')
	return(out)	
}
```

**Mixed model for lefts**
```{r}
LModel.lmer		= lmer(Z ~ 1 + subjId + BinNum + ROI + (1 + BinNum|chanId) , data=data , subset=subL , REML = F)
LModelInt.lmer	= lmer(Z ~ 1 + subjId + BinNum * ROI + (1 + BinNum|chanId) , data=data , subset=subL , REML = F)
LRT(LModel.lmer,3)
LRT(LModelInt.lmer,4)
```

**Mixed model for rights**
```{r}
RModel.lmer	    = lmer(Z ~ 1 + subjId + BinNum + ROI + (1 + BinNum|chanId) , data=data , subset=subR , REML = F)
RModelInt.lmer	= lmer(Z ~ 1 + subjId + BinNum * ROI + (1 + BinNum|chanId) , data=data , subset=subR , REML = F)
LRT(RModel.lmer,3)
LRT(RModelInt.lmer,4)
```

**Mixed model to test hemisphere interaction.**
Note that the use of subject as a factor in this model results in an 
overdetermined matrix. 
```{r}
FModel.lmer  	= lmer(Z ~ 1 + BinNum * ROI * hem + (1 + BinNum|chanId) , data=data , REML = F)
LRT(FModel.lmer,7)
FModel.lmer  	= lmer(Z ~ 1 + subjId + BinNum * ROI * hem + (1 + BinNum|chanId) , data=data , REML = F)
```

